---
title: "Position Paper"
format: 
  html:
    toc: true
    toc-depth: 5
    grid:
      sidebar-width: 300px
      body-width: 2000px
      margin-width: 300px
      gutter-width: 1.5rem
editor: visual
execute:
  warning: false
---

## Packages

```{r}
# lapply(paste0('package:', names(sessionInfo()$otherPkgs)), detach, character.only=TRUE, unload=TRUE)

library(dplyr)
library(readxl)
library(tidyr)
library(purrr)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(stringr)
library(paletteer)
```

## Setup

```{r, setup}
# knitr::opts_knit$set(root.dir = "C:/Users/rfrost/OneDrive - The Institute of Cancer Research/Documents/Position Paper/Data")
knitr::opts_knit$set(root.dir = "C:/Users/rfrost/OneDrive - The Institute of Cancer Research/Documents/UK-cancer-trends/Data")
knitr::opts_chunk$set(echo = FALSE)
theme_set(theme_minimal())
```

## Function - Main

```{r}
## Main function that grabs data depending on source and calls specific cleaning function,
data_hub <- function(country, rf){
  
  # country <- "Wales"
  # 
  # rf <- "E_Cigarette"
  
  data <- list()
  
  sheets <- excel_sheets(paste0(rf, "_Data.xlsx"))
  
  spec_sheets <- sheets[grepl(sheets, pattern = country)]
  
  if(country == "England"){
    for(sheet in spec_sheets){
      nocol <- ncol(read_excel(paste0(rf, "_Data.xlsx"), sheet = sheet))
      tempdata <- read_excel(paste0(rf, "_Data.xlsx"), sheet = sheet, col_types = c("text", rep("numeric", nocol-1)))
      group <- gsub(x = sheet, pattern = "England_", replacement = "")
      data[[paste0(rf, "-", country, "-", sheet)]] <- pull_HSE_data(data = tempdata, group = group)
      data[[paste0(rf, "-", country, "-", sheet)]]$section <- rf
    }
  } else if(country == "Scotland"){
    for(sheet in spec_sheets){
      tempdata <- read_excel(paste0(rf, "_Data.xlsx"), sheet = sheet)
      group <- gsub(x = sheet, pattern = "Scotland_", replacement = "")
      data[[paste0(rf, "-", country, "-", sheet)]] <- pull_SHS_data(data = tempdata, group = group)
          data[[paste0(rf, "-", country, "-", sheet)]]$section <- rf
    }
  } else if(country == "Wales"){
    for(sheet in spec_sheets){
      tempdata <- read_excel(paste0(rf, "_Data.xlsx"), sheet = sheet)
      group <- gsub(x = sheet, pattern = "Wales_", replacement = "")
      data[[paste0(rf, "-", country, "-", sheet)]] <- pull_NSW_data(data = tempdata, group = group)
      data[[paste0(rf, "-", country, "-", sheet)]]$section <- rf

    }
  } else if(country == "NIreland" & rf != "Activity"){
    for(sheet in spec_sheets){
      tempdata <- read_excel(paste0(rf, "_Data.xlsx"), sheet = sheet)
      group <- gsub(x = sheet, pattern = "NIreland_", replacement = "")
      data[[paste0(rf, "-", country, "-", sheet)]] <- pull_HSNI_data(data = tempdata, group = group)
      data[[paste0(rf, "-", country, "-", sheet)]]$variable <- get_variable_HSNI(rf)
      data[[paste0(rf, "-", country, "-", sheet)]]$section <- rf
    }
    # Activity data formatted differently than other N. Ireland data so need different formatting.
  } else if(country == "NIreland" & rf == "Activity"){
    
    tempdata <- read_excel(paste0(rf, "_Data.xlsx"), sheet = "NIreland_Men")
    group <- "Men"
    data[["Activity-NIreland-Men"]] <- pull_HSNI_data_alt(data = tempdata, group = "Men")
    # data[["Activity-NIreland-Men"]]$variable <- get_variable_HSNI(rf)
    data[["Activity-NIreland-Men"]]$section <- rf
    
    tempdata <- read_excel(paste0(rf, "_Data.xlsx"), sheet = "NIreland_Women")
    group <- "Women"
    data[["Activity-NIreland-Women"]] <- pull_HSNI_data_alt(data = tempdata, group = "Women")
    # data[["Activity-NIreland-Women"]]$variable <- get_variable_HSNI(rf)
    data[["Activity-NIreland-Women"]]$section <- rf

    
    tempdata <- read_excel(paste0(rf, "_Data.xlsx"), sheet = "NIreland_All")
    group <- "All"
    data[["Activity-NIreland-All"]] <- pull_HSNI_data(data = tempdata, group = "All")
    data[["Activity-NIreland-All"]]$variable <- get_variable_HSNI(rf)
    data[["Activity-NIreland-All"]]$section <- rf
    
  } else {
    return(cat("ERROR"))
  }
  
  return(data)
}

```

## Function - England Health Survey

```{r}

pull_HSE_data <- function(data, group){
  
  # Edit for activity data format
  if(is.na(data[2,2])){
    data <- data[-1,]
  }
  
  # Remove weighted bases - can alter to keep
  if("Weighted bases" %in% data$...1){
    data <- data[1:(which(data$`...1`=="Weighted bases")-1),]
  } else if("Bases (weighted)" %in% data$...1){
    data <- data[1:(which(data$`...1`=="Bases (weighted)")-1),]
  }
  
  # Get cat columns
  ranges <- data$...1[is.na(data[,2])]
  rangesdif <- diff(which(is.na(data[,2])))
  
  if(max(rangesdif) != min(rangesdif)){
    return(cat("ERROR: Inconsistent Data Format"))
  }
  
  data <- data |>
    mutate(age_group = rep(ranges, each = rangesdif[1])) |>
    filter(!`...1` %in% ranges) |>
    pivot_longer(cols = colnames(data)[2:ncol(data)]) |>
    rename(year = name, variable = `...1`)
  
  data[["country"]] <- "England"
  data[["group"]] <- group
  
  ## Smoking data for some reason is ordered oppositely, so made quick fix here.
  ## Not ideal fix but will do for temporary purposes.
  if("Current cigarette smoker" %in% data[["age_group"]] | "Currently uses e-cigarettes" %in% data[["age_group"]] ){
    data <- data |>
      rename(temp_age = variable, temp_var = age_group) |>
      rename(age_group = temp_age, variable = temp_var)
  }
  
  return(data)
  
}

```

## Function - Alt England Health Survey (Diabetes Data)

```{r}

pull_HSE_diabetes_data <- function(data, group){
  
  if("Weighted bases" %in% data$...1){
    data <- data[1:(which(data$`...1`=="Weighted bases")-1),]
  }
  
  colnames(data) <- c("age_group", colnames(data)[2:ncol(data)])
  
  data <- data[-c(1,2),]
  
  data <- data |>
    pivot_longer(cols = colnames(data)[2:ncol(data)]) |>
    rename(year = name)
  
  data[["country"]] <- "England"
  data[["group"]] <- group
  data[["variable"]] <- "% Diabetes"
  
  return(data)
  
}

```

## Function - Scottish Health Survey

```{r}

pull_SHS_data <- function(data, group){
  
  if(group == "Sex"){
    data <- data |>
      select(year = Year, country = Geographylevel, variable = Categories, 
             group = Sex, value = Percent) |>
      mutate(age_group = "16+")
  } else if(group == "Age"){
    data <- data |>
      select(year = Year, country = Geographylevel, 
             variable = Categories, value = Percent, age_group = Age) |>
      mutate(group = "All adults")
  }
  
  return(data)
  
}
```

## Function - National Survey for Wales

```{r}

pull_NSW_data <- function(data, group){

  year_input <- group

  # data <- read_excel("E_Cigarette_Data.xlsx", sheet = "Wales_2017_2018")
  
  data <- as.data.frame(t(as.matrix(data, dimnames = NULL)))
  
  rownames(data) <- NULL
  
  data[data=="."] <- NA
  
  data <- data |>
    mutate(V2 = case_when(
      is.na(V2) ~ V1,
      TRUE ~ V2
    ), V3 = case_when(
      is.na(V3) ~ V2,
      TRUE ~ V3
    )) |>
    select(-V1, -V2)
  
  colnames(data) <- c("age_group", data[2,2:ncol(data)])
  
  data <- data[-c(1,2),]
  
  
  data <- data |>
    mutate(group = rep(c("All", "Men", "Women"), each = 11),
           year = year_input,
           country = "Wales") |>
    pivot_longer(cols = colnames(data)[2:ncol(data)]) |>
    rename(variable = name)
  
  return(data)
}
```

## Function - Health Survey Northern Ireland

```{r}
pull_HSNI_data <- function(data, group){
  
  # data <- read_excel("Activity_Data.xlsx", sheet = "NIreland_Men")

  # group <- "Males"
  
  colnames(data)[1] <- "age_group"
  
  data <-  data |>
    pivot_longer(cols = colnames(data)[2:ncol(data)]) |>
    rename(year = name) |>
    mutate(country = "N. Ireland",
           group = {{group}})
  
  return(data)
}

pull_HSNI_data_alt <- function(data, group){
  
  # data <- read_excel("Activity_Data.xlsx", sheet = "NIreland_Men")

  # group <- "Males"
  
  colnames(data)[1] <- "variable"
  
  data <-  data |>
    pivot_longer(cols = colnames(data)[2:ncol(data)]) |>
    rename(year = name) |>
    mutate(country = "N. Ireland",
           group = {{group}},
           age_group = "19+") # Table specifies 19+ participants.
  
  return(data)
}

get_variable_HSNI <- function(rf){
  section <- rf
  variable <- case_when(
    section == "BMI" ~ "Overweight",
    section == "Alcohol" ~ "Drink",
    section == "Smoking" ~ "Smoke",
    section == "E_Cigarette" ~ "E-Cigarette Use",
    section == "Activity" ~ "Meet Recommendations",
    TRUE ~ "ERROR"
  )
}
```

## Grabbing First Round of Data

Grabbing data where all sources are similar.

```{r}
risk_factors <- c("BMI", "Alcohol", "E_Cigarette", "Smoking", "Activity")
countries <- c("England", "Scotland", "Wales", "NIreland")

data <- data.frame(section = character(0), variable = character(0), country = character(0), year = numeric(0),
                   group = character(0), age_group = character(0), value = numeric(0))

for(rf in risk_factors){
  for(country in countries){
    new_data <- bind_rows(data_hub(country, rf))
    data <- rbind(data, new_data)
  }
}


```

## Cleaning Data

Cleaning some specific variables as well as cleaning variable names for variables needed in analysis.

```{r}

data2 <- data |>
  mutate(year = case_when(
    # Choosing first year when there are two 2017/18, may chance depending on future approach.
    grepl(year, pattern = "\\/") ~ map_chr(strsplit(year, "/"), 1),
    grepl(year, pattern = "\\_") ~ map_chr(strsplit(year, "_"), 1),
    TRUE ~ year
  ), 
  year = as.numeric(year),
  age_group = case_when(
    age_group %in% c("All men", "All women", "All", "All adults", "Total") ~ "16+",
    grepl(age_group, pattern = "Age ") ~ gsub(age_group, pattern = "Age ", replacement = ""),
    grepl(age_group, pattern = " years") ~ gsub(age_group, pattern = " years", replacement = ""),
    grepl(age_group, pattern = "All adults ") ~ gsub(age_group, pattern = "All adults ", replacement = ""),
    grepl(age_group, pattern = "All aged ") ~ gsub(age_group, pattern = "All aged ", replacement = ""),
    TRUE ~ age_group
  ), age_group = case_when(
    # Expections to all being 16+
    country == "N. Ireland" & section == "Alcohol" ~ gsub(age_group, pattern = "16", replacement = "18"),
    country == "N. Ireland" & section == "Activity" ~ gsub(age_group, pattern = "16", replacement = "19"),
    TRUE ~ age_group
  ), variable = case_when(
    # Harmonising variable names. Only doing ones we are focussed on. Could always expand to a fully harmonised dataset.
    country == "England" & section == "BMI" & variable == "Overweight including obese" & grepl(group, pattern = "Children") ~ "% Overweight",
    country == "England" & section == "BMI" & variable == "Obese" & grepl(group, pattern = "Children") ~ "% Obese",
    country == "England" & section == "BMI" & variable == "% overweight including obese" & !grepl(group, pattern = "Children") ~ "% Overweight",
    country == "England" & section == "BMI" & variable == "% obese" & !grepl(group, pattern = "Children") ~ "% Obese",
    country == "England" & section == "BMI" & variable == "% overweight" ~ "% overweight not obese",
    country == "England" & section == "Alcohol" & variable == "% non drinker/did not drink in last 12 months" ~ "% Non-Drinker",
    country == "England" & section == "Smoking" & variable == "Current cigarette smoker" ~ "% Current Smoker",
    country == "England" & section == "E_Cigarette" & variable == "Currently uses e-cigarettes" ~ "% Current E-cigarette user",
    country == "England" & section == "Activity" & variable %in% c("Meets aerobic guidelines", "Meets previous aerobic guidelines") ~ "% Meets guidelines", ## CHECK THIS IS APPROPRIATE
    ## Will probably use the more dense activity dataset for England so this ceases to be an issue in that case.
    
    country == "Scotland" & section == "BMI" & variable == "Overweight (including obesity)" ~ "% Overweight",
    country == "Scotland" & section == "Alcohol" & variable == "Non-drinker" ~ "% Non-Drinker",
    country == "Scotland" & section == "Smoking" & variable == "Current smoker" ~ "% Current Smoker",
    country == "Scotland" & section == "E_Cigarette" & variable == "Currently using" ~ "% Current E-cigarette user",
    country == "Scotland" & section == "Activity" & variable == "Meets recommendations" ~ "% Meets guidelines",
    
    country == "Wales" & section == "BMI" & variable == "BMI 25+ (overweight or obese)" ~ "% Overweight",
    country == "Wales" & section == "BMI" & variable == "BMI 30+ (obese)" ~ "% Obese",
    country == "Wales" & section == "Alcohol" & variable == "Average weekly alcohol consumption - none (may include some people who do sometimes drink)" ~ "% Non-Drinker",
    country == "Wales" & section == "Smoking" & variable == "Smoker" ~ "% Current Smoker",
    country == "Wales" & section == "E_Cigarette" & variable == "E-cigarette user" ~ "% Current E-cigarette user",
    country == "Wales" & section == "Activity" & variable == "Active at least 150 minutes in previous week" ~ "% Meets guidelines",
    
    country == "N. Ireland" & section == "BMI" & variable == "Overweight" ~ "% Overweight",
    country == "N. Ireland" & section == "Alcohol" & variable == "Drink" ~ "% Non-Drinker", # NOTE: This is not technically accurate but next code block rectifies.
    country == "N. Ireland" & section == "Smoking" & variable == "Smoke" ~ "% Current Smoker",
    country == "N. Ireland" & section == "E_Cigarette" & variable == "E-Cigarette Use" ~ "% Current E-cigarette user",
    country == "N. Ireland" & section == "Activity" & variable == "Meet Recommendations" ~ "% Meets guidelines",
    
    TRUE ~ variable

  ))

data3 <-  data2 |>
  mutate(value = as.numeric(value),
         value = case_when(
    country == "N. Ireland" & section == "Alcohol" & variable == "% Non-Drinker" ~ 1 - value,
    TRUE ~ value
  ), value = case_when(
    country == "N. Ireland" ~ 100 * value,
    TRUE ~ value
  ))

## Need to group England overweight values together.

# england_overweight <- data3 |>
#   filter(country == "England" & section == "BMI" & !grepl(group, pattern = "Children")) |>
#   pivot_wider(names_from = variable, values_from = value) |>
#   mutate(`% Overweight` = `% overweight including obese`) |>
#   pivot_longer(cols = c(unique(data3$variable[data3$country == "England" & data3$section == "BMI" & !grepl(data3$group, pattern = "Children")]), "% Overweight")) |>
#   rename(variable = name)

# data4 <- data3 |>
#   # filter(!(country == "England" & section == "BMI" & !grepl(group, pattern = "Children"))) |>
#   mutate(variable = case_when(
#     country == "England" & section == "BMI" & variable == "Overweight including obese" & grepl(group, pattern = "Children") ~ "% Overweight",
#     country == "England" & section == "BMI" & variable == "% overweight including obese" & !grepl(group, pattern = "Children") ~ "% Overweight",
#     TRUE ~ variable
#   )) # |>
  # rbind(england_overweight)

data4 <- data3

```

## Overweight

```{r}

data4 |>
  filter(variable == "% Overweight" & age_group == "16+" & grepl(group, pattern = "All")) |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Overweight") +
  ggtitle("% Overweight in Adults") +
  labs(colour =  "Country") +
  scale_x_continuous(breaks = pretty_breaks())

```

## Obese

```{r}
obese_data <- data4

obese_sheets <- excel_sheets("BMI_Obese_Data.xlsx")

scotland_sheets <- obese_sheets[grepl(obese_sheets, pattern = "Scotland")]

for(sheet in scotland_sheets){
  tempdata <- read_excel("BMI_Obese_Data.xlsx", sheet = sheet)
  group <- gsub(x = sheet, pattern = "Scotland_", replacement = "")
  tempdata <- pull_SHS_data(data = tempdata, group = group)
  tempdata$section <- "BMI"
  tempdata$variable <- ifelse(tempdata$variable == "Obesity", "% Obese", "% Non-Obese")
  
  
  obese_data <- rbind(obese_data, tempdata)

}

nireland_sheets <- obese_sheets[grepl(obese_sheets, pattern = "NIreland")]

for(sheet in nireland_sheets){
  tempdata <- read_excel(paste0("BMI_Obese_Data.xlsx"), sheet = sheet)
  group <- gsub(x = sheet, pattern = "NIreland_", replacement = "")
  tempdata <- pull_HSNI_data(data = tempdata, group = group)
  tempdata$variable <- "% Obese"
  tempdata$section <- "BMI"
  tempdata$value <- 100*tempdata$value
  
  obese_data <- rbind(obese_data, tempdata)
}

obese_data |>
  filter(variable == "% Obese" & age_group %in% c("16+", "Total") & grepl(group, pattern = "All")) |>
  mutate(year = case_when(
  grepl(year, pattern = "\\/") ~ as.numeric(map_chr(strsplit(year, "/"), 1)),
  TRUE ~ as.numeric(year)
  )) |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Obese") +
  ggtitle("% Obese in Adults") +
  labs(colour =  "Country") +
  scale_x_continuous(breaks = pretty_breaks())


```

## Obese by Age - England

```{r}
# Need to use weightings from the data set to calculate % Obese for 25+ age range.
weightings <- read_excel("BMI_Data.xlsx", sheet = "England_All")

top_row <- which(weightings$...1 == "Weighted bases")
weightings <-weightings[(top_row+1):nrow(weightings),]

weightings <- as.data.frame(weightings)


weightings <- weightings |>
  filter(!...1 %in% c("All adults", "16-24 years"))

for(i in 2:ncol(weightings)){
  weightings[,i] <- weightings[,i]/sum(weightings[,i])
}

weightings <- as.data.frame(t(as.matrix(weightings)))

weightings <- weightings |>
  mutate(Year = rownames(weightings))

rownames(weightings) <- NULL
colnames(weightings) <- weightings[1,]
weightings <- weightings[-1,]

weightings <- weightings |>
  rename(year = ...1) |>
  sapply(as.numeric)

adult_bmi_england <- data4 |>
  filter(country == "England" & section == "BMI" & variable == "% Obese" & group == "All" & !grepl(age_group, pattern = "16")) |>
  pivot_wider(names_from = age_group, values_from = value) |>
  merge(weightings, by = "year") |>
  mutate(`25+` = `25-34`*`25-34 years` + `35-44`*`35-44 years` + `45-54`*`45-54 years` + `55-64`*`55-64 years` + `65-74`*`65-74 years` + `75+`*`75+ years`) |>
  select(year, variable, country, group, section, `25+`) |>
  pivot_longer(cols = `25+`) |>
  rename(age_group = name)

age_bmi_england <- data4 |>
  filter(country == "England" & section == "BMI" & variable == "% Obese" & group %in% c("All", "Children_All") & age_group %in% c("16-24", "2-10", "11-15")) |>
  rbind(adult_bmi_england) 

ggplot(data = age_bmi_england, aes(x = year, y = value, colour = factor(age_group, levels = c("2-10", "11-15", "16-24", "25+")))) +
  geom_point() +
  geom_line() +
  xlab("Year") +
  ylab("% Obese") +
  ggtitle("% Obese in England") +
  labs(colour =  "Age Group", caption = "*Obese for those aged >15 as BMI > 30\nObese for those aged <= 15 as at or above the 95th BMI percentile in the UK") +
  scale_x_continuous(breaks = pretty_breaks())
 

```

## Childhood Obesity - NCMP
Childhood obesity as measured by the National Child Measurement Program
```{r}

reception_BMI_data <- read_excel("BMI_Data_Children.xlsx", sheet = "England_Reception")

reception_BMI_data <- as.data.frame(t(as.matrix(reception_BMI_data)))

colnames(reception_BMI_data) <- c("variable", "var2", "REMOVE", reception_BMI_data[1,4:ncol(reception_BMI_data)])
reception_BMI_data <- reception_BMI_data |>
  filter(row_number()>1 & row_number()<nrow(reception_BMI_data)) |>
  mutate(variable = rep(reception_BMI_data$variable[!is.na(reception_BMI_data$variable)], each = 5)) |>
  filter(var2 == "Prevalence") |>
  select(-REMOVE, -var2, -Reception) |>
  pivot_longer(cols = colnames(reception_BMI_data)[grepl(colnames(reception_BMI_data), pattern = "20")]) |>
  mutate(year = as.numeric(substr(name, start = 1, stop = 4)),
         age_group = "4-5")

yr6_BMI_data <- read_excel("BMI_Data_Children.xlsx", sheet = "England_Year_6")

yr6_BMI_data <- as.data.frame(t(as.matrix(yr6_BMI_data)))

colnames(yr6_BMI_data) <- c("variable", "var2", "REMOVE", yr6_BMI_data[1,4:ncol(yr6_BMI_data)])
yr6_BMI_data <- yr6_BMI_data |>
  filter(row_number()>1 & row_number()<nrow(yr6_BMI_data)) |>
  mutate(variable = rep(yr6_BMI_data$variable[!is.na(yr6_BMI_data$variable)], each = 5)) |>
  filter(var2 == "Prevalence") |>
  select(-REMOVE, -var2, -`Year 6`) |>
  pivot_longer(cols = colnames(yr6_BMI_data)[grepl(colnames(yr6_BMI_data), pattern = "20")]) |>
  mutate(year = as.numeric(substr(name, start = 1, stop = 4)),
         age_group = "10-11")

children_BMI_data <- rbind(reception_BMI_data, yr6_BMI_data)

children_BMI_data |>
  filter(variable == "Obesity (including severe obesity)") |>
  filter(year != 2020) |> # They noted 2020 as having a smaller sample and needing weighting so we have excluded it here.
  mutate(value = as.numeric(value)) |>
  ggplot(aes(x = year, y = value, colour = age_group)) +
  geom_point() +
  geom_line() +
  xlab("Year") +
  ylab("% Obese") +
  ggtitle("% Obese in England  - Children") +
  labs(colour =  "Age Group") +
  scale_x_continuous(breaks = pretty_breaks())
  

```

## Alcohol

```{r}

data4 |>
  filter(variable == "% Non-Drinker" & age_group %in% c("16+", "18+") & grepl(group, pattern = "All")) |>
  mutate(value = 100- value) |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Current Drinker") +
  ggtitle("% Current Drinker in Adults*") +
  labs(colour =  "Country", caption = "*Adult for England, Scotland and Wales defined as 16+, and for N. Ireland as 18+") +
  scale_x_continuous(breaks = pretty_breaks())


```

## Alcohol - Guidelines

Needed to get guidelines data for N. Ireland seperately, as well as combine variables for England data to get a Drinking over Guidelines value.

```{r}

guidelines_england <- data4 |>
  filter(country == "England" & section == "Alcohol"& grepl(group, pattern = "All")) |>
  pivot_wider(names_from = variable, values_from = value) |>
  mutate(`% Over Guidelines` = 100 - (`% Non-Drinker` + `% up to 14 units per week`)) |>
  pivot_longer(cols = c(unique(data4$variable[data4$country == "England" & data4$section == "Alcohol" & grepl(data4$group, pattern = "All")]), "% Over Guidelines")) |>
  rename(variable = name)

alcohol_data <- data4 |>
  filter(!(country == "England" & section == "Alcohol"& grepl(group, pattern = "All"))) |>
  rbind(guidelines_england) |>
  mutate(variable = case_when(
    country == "Scotland" & section == "Alcohol" & variable == "Hazardous/Harmful drinker" ~ "% Over Guidelines",
    country == "Wales" & section == "Alcohol" & variable == "Average weekly alcohol consumption - above 14 units (over guidelines)" ~ "% Over Guidelines",
    TRUE ~ variable
  ))

new_alcohol_sheets <- excel_sheets("Alcohol_Data_Alt.xlsx")
for(sheet in new_alcohol_sheets){
  tempdata <- read_excel(paste0("Alcohol_Data_Alt.xlsx"), sheet = sheet)
  group <- gsub(x = sheet, pattern = "NIreland_", replacement = "")
  tempdata <- pull_HSNI_data(data = tempdata, group = group)
  tempdata$variable <- "% Over Guidelines"
  tempdata$section <- "Alcohol"
  tempdata$value <- 100*tempdata$value
  
  alcohol_data <- rbind(alcohol_data, tempdata)
}

alcohol_data |>
  filter(variable == "% Over Guidelines" & group == "All" & age_group %in% c("16+", "Total")) |>
  mutate(year = case_when(
    grepl(year, pattern = "\\/") ~ map_chr(strsplit(year, "/"), 1),
    TRUE ~ year
  )) |>
  ggplot(aes(x = as.numeric(year), y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Over Alcohol Guidelines") +
  ggtitle("% Over Alcohol Guidelines in Adults") +
  labs(colour =  "Country") +
  scale_x_continuous(breaks = pretty_breaks())
```

## Alcohol - Guidelines by Age

```{r}

guidelines_england <- data4 |>
  filter(country == "England" & section == "Alcohol"& grepl(group, pattern = "All")) |>
  pivot_wider(names_from = variable, values_from = value) |>
  mutate(`% Over Guidelines` = 100 - (`% Non-Drinker` + `% up to 14 units per week`)) |>
  pivot_longer(cols = c(unique(data4$variable[data4$country == "England" & data4$section == "Alcohol" & grepl(data4$group, pattern = "All")]), "% Over Guidelines")) |>
  rename(variable = name)

# Need to use weightings from the data set to calculate % Obese for 25+ age range.
weightings <- read_excel("Alcohol_Data.xlsx", sheet = "England_All")

top_row <- which(weightings$...1 == "Weighted bases")
weightings <-weightings[(top_row+1):nrow(weightings),]

weightings <- as.data.frame(weightings)


weightings <- weightings |>
  filter(!...1 %in% c("All adults", "16-24 years"))

for(i in 2:ncol(weightings)){
  weightings[,i] <- weightings[,i]/sum(weightings[,i])
}

weightings <- as.data.frame(t(as.matrix(weightings)))

weightings <- weightings |>
  mutate(Year = rownames(weightings))

rownames(weightings) <- NULL
colnames(weightings) <- weightings[1,]
weightings <- weightings[-1,]

weightings <- weightings |>
  rename(year = ...1) |>
  sapply(as.numeric)

adult_alcohol_england <- guidelines_england |>
  filter(variable == "% Over Guidelines" & !grepl(age_group, pattern = "16")) |>
  pivot_wider(names_from = age_group, values_from = value) |>
  merge(weightings, by = "year") |>
  mutate(`25+` = `25-34`*`25-34 years` + `35-44`*`35-44 years` + `45-54`*`45-54 years` + `55-64`*`55-64 years` + `65-74`*`65-74 years` + `75+`*`75+ years`) |>
  select(year, variable, country, group, section, `25+`) |>
  pivot_longer(cols = `25+`) |>
  rename(age_group = name)

age_alcohol_england <- guidelines_england |>
  filter(variable == "% Over Guidelines" & age_group %in% c("16-24")) |>
  rbind(adult_alcohol_england)

ggplot(data = age_alcohol_england, aes(x = year, y = value, colour = age_group)) +
  geom_point() +
  geom_line() +
  xlab("Year") +
  ylab("% Over Guidelines") +
  ggtitle("% Over Guidelines in England") +
  labs(colour =  "Age Group") +
  scale_x_continuous(breaks = pretty_breaks())

```

## Smoking

```{r}

data4 |>
  filter(variable == "% Current Smoker" & age_group == "16+" & grepl(group, pattern = "All")) |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Current Smoker") +
  ggtitle("% Current Smoker in Adults") +
  labs(colour =  "Country") +
  scale_x_continuous(breaks = pretty_breaks())


# data4 |>
#   filter(variable == "% Current Smoker" & age_group == "16+" & !grepl(group, pattern = "All")) |>
#   mutate(group = case_when(
#     group == "Male" ~ "Men",
#     group == "Female" ~ "Women",
#     TRUE ~ group
#   )) |>
#   ggplot(aes(x = year, y = value, colour = country, linetype = group)) +
#   geom_line() +
#   geom_point() +
#   xlab("Year") +
#   ylab("% Current Smoker") +
#   ggtitle("% Current Smoker in Adults") +
#   labs(colour =  "Country") +
#   scale_x_continuous(breaks = pretty_breaks())

data4 |>
  filter(variable == "% Current Smoker" & age_group == "16+" & !grepl(group, pattern = "All")) |>
  mutate(group = case_when(
    group == "Male" ~ "Men",
    group == "Female" ~ "Women",
    TRUE ~ group
  )) |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Current Smoker") +
  ggtitle("% Current Smoker in Adults") +
  labs(colour =  "Country") +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_wrap(~group, scales = "fixed", dir = "v")

```

## Smoking by Age

```{r}

weightings <- read_excel("Smoking_Data.xlsx", sheet = "England_All")

top_row <- which(weightings$...1 == "Weighted bases")
weightings <-weightings[(top_row+1):nrow(weightings),]

weightings <- as.data.frame(weightings)


weightings <- weightings |>
  filter(!...1 %in% c("All adults", "16-24 years"))

for(i in 2:ncol(weightings)){
  weightings[,i] <- weightings[,i]/sum(weightings[,i])
}

weightings <- as.data.frame(t(as.matrix(weightings)))

weightings <- weightings |>
  mutate(Year = rownames(weightings))

rownames(weightings) <- NULL
colnames(weightings) <- weightings[1,]
weightings <- weightings[-1,]

weightings <- weightings |>
  rename(year = ...1) |>
  sapply(as.numeric)

adult_smoking_england <- data4 |>
  filter(variable == "% Current Smoker" & !grepl(age_group, pattern = "16") & country == "England" & group == "All") |>
  pivot_wider(names_from = age_group, values_from = value) |>
  merge(weightings, by = "year") |>
  mutate(`25+` = `25-34`*`25-34 years` + `35-44`*`35-44 years` + `45-54`*`45-54 years` + `55-64`*`55-64 years` + `65-74`*`65-74 years` + `75+`*`75+ years`) |>
  select(year, variable, country, group, section, `25+`) |>
  pivot_longer(cols = `25+`) |>
  rename(age_group = name)

age_smoking_england <- data4 |>
  filter(variable == "% Current Smoker" & country == "England" & group == "All" & age_group %in% c("16-24")) |>
  rbind(adult_smoking_england)


ggplot(data = age_smoking_england, aes(x = year, y = value, colour = age_group)) +
  geom_point() +
  geom_line() +
  xlab("Year") +
  ylab("% Current Smoker") +
  ggtitle("% Current Smoker in England") +
  labs(colour =  "Age Group") +
  scale_x_continuous(breaks = pretty_breaks())

```

## E-Cigarette

```{r}

data4 |>
  filter(variable == "% Current E-cigarette user" & age_group == "16+" & grepl(group, pattern = "All")) |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Current E-Cigarette Users") +
  ggtitle("% Current E-Cigarette Users in Adults") +
  labs(colour =  "Country") +
  scale_x_continuous(breaks = pretty_breaks())

cols <- colorRampPalette(brewer.pal(n = 8, name = "Greens")[4:8])(7)

data4 |>
  filter(variable == "% Current E-cigarette user" & grepl(group, pattern = "All") & country == "England" & age_group != "16+") |>
  mutate(age_start = as.numeric(substr(age_group, 1, 2)), 
         age_end = ifelse(grepl(age_group, pattern = "\\+"), age_start, as.numeric(substr(age_group, 4, 5))),
         age_use = 0.5*(age_start + age_end)) |>
  ggplot(aes(x = year, y = value, colour = as.factor(age_group))) +
  geom_line() +
  geom_point() +
  xlab("Age") +
  ylab("% Current E-Cigarette Users") +
  ggtitle("% Current E-Cigarette Users in the UK") +
  labs(colour =  "Age Group") +
  scale_x_continuous(breaks = 2016:2019) +
  scale_color_manual(values = cols)
```

## E-Cigarette - Great Britain

More years available - but concerns about different questionnaire structure.

```{r}

ecig_data <- read_excel("E_Cigarette_Data.xlsx", sheet = "Great_Britain_All")

ecig_data <- ecig_data |>
  pivot_longer(cols = colnames(ecig_data)[3:ncol(ecig_data)]) |>
  pivot_wider(names_from = `Proportion of population aged 16 years and over by e-cigarette use [note 1]`, values_from = value) |>
  mutate(`% User` = `Daily user (%)` + `Occasional user (%)`) |>
  select(Year, name, `% User`) |>
  separate(col = name, sep = " \r\naged \r\n", into = c("group", "age_group")) |>
  mutate(Year = as.integer(Year), 
         age_group = case_when(
           grepl(age_group, pattern = " to ") ~ gsub(age_group, pattern = " to ", replacement = "-"),
           grepl(age_group, pattern = " and over") ~ gsub(age_group, pattern = " and over", replacement = "+"),
           TRUE ~ age_group
         ))

ecig_data |>
  filter(group == "All persons", age_group == "16+") |>
  ggplot(aes(x = Year, y = `% User`)) +
  geom_line() +
  geom_point() +
  ylab("% Current E-Cigarette Users") +
  ggtitle("% Current E-Cigarette Users in Great Britain") +
  scale_x_continuous(breaks = pretty_breaks())

cols <- brewer.pal(n = 8, name = "Greens")[4:8]

ecig_data |>
  filter(!age_group %in% c("16+"), group == "All persons") |>
  ggplot(aes(x = Year, y = `% User`, colour = age_group)) +
  geom_line() +
  geom_point() +
  ylab("% Current E-Cigarette Users") +
  ggtitle("% Current E-Cigarette Users in Great Britain") +
  labs(colour = "Age Group") +
  scale_color_manual(values = cols) +
  scale_x_continuous(breaks = ecig_data$Year[ecig_data$Year%%1==0])

```

## Activity

Added a new source for the activity that provides more years.
The Welsh definition is slightly differently worded, check up on this.
```{r}

data4 |>
  filter(variable == "% Meets guidelines" & age_group %in% c("16+") & grepl(group, pattern = "All") |
           variable == "% Meets guidelines" & age_group %in% c("19+") & grepl(group, pattern = "All") & country == "N. Ireland") |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Meets Activity Guidelines") +
  ggtitle("% Meets Activity Guidelines in Adults*") +
  labs(colour =  "Country", caption = "*Adult for England, Scotland and Wales defined as 16+, and for N. Ireland as 19+\nActivity guidelines are >150 mins Moderate Physical Activity") +
  scale_x_continuous(breaks = pretty_breaks())

# Adding alternative data source for England

activity_england <- read_excel("Activity_Alt_Data.xlsx")

activity_england <- activity_england |>
  rename(variable = ...1) |>
  pivot_longer(cols = colnames(activity_england[2:ncol(activity_england)])) |>
  rename(year = name) |>
  mutate(year = gsub(year, pattern = "Nov ", replacement = "")) |>
  mutate(year = as.numeric(paste0("20", substr(year, start = 1, stop = 2))),
         value = 100*as.numeric(value),
         variable = case_when(
           variable == "Active: at least 150 minutes a week" ~ "% Meets guidelines",
           TRUE ~ variable
         ), age_group = "16+",
         group = "All",
         country = "England",
         section = "Activity") |>
  filter(variable == "% Meets guidelines")

data4 |>
  filter(variable == "% Meets guidelines" & age_group %in% c("16+") & grepl(group, pattern = "All") & country != "England" |
           variable == "% Meets guidelines" & age_group %in% c("19+") & grepl(group, pattern = "All") & country == "N. Ireland") |>
  rbind(activity_england) |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Meets Activity Guidelines") +
  ggtitle("% Meets Activity Guidelines in Adults*") +
  labs(colour =  "Country", caption = "*Adult for England, Scotland and Wales defined as 16+, and for N. Ireland as 19+\nActivity guidelines are >150 mins Moderate Physical Activity") +
  scale_x_continuous(breaks = pretty_breaks())

```

## Age First Birth Data

For this variable and all further, no data is contained in the pre sorted data, so all sourced from new sources. As with before, these sources have been documented in the README file.

```{r}

afb_data <- read_excel("Age_First_Birth_Data.xlsx")[2:3,13:31]

colnames(afb_data) <- afb_data[1,]
afb_data <- afb_data[2,]

afb_data |>
  pivot_longer(cols = colnames(afb_data)) |>
  mutate(Year = as.numeric(name), value = as.numeric(value)) |>
  ggplot(aes(x = Year, y = value)) +
  geom_line() +
  geom_point() +
  ylab("Average Age at First Birth") +
  ggtitle("Average Age at First Birth in UK") +
  scale_x_continuous(breaks = pretty_breaks())

```

## Parity

Currently only have data for England and Wales. Trying to track down data for other countries.

```{r}
# 
# read_excel("Parity_Data.xlsx") |>
#   mutate(age = case_when(
#     is.na(`Age exact years`) ~ 45,
#     TRUE ~ `Age exact years`),
#     year = age + `Year of birth of woman`,
#     final_value = ifelse(is.na(`Age exact years`), 1, 0),
#     `% Parous` = 100 - `0 (No children)`) |>
#   filter(final_value == 1) |>
#   ggplot(aes(x = year, y = `% Parous`)) +
#   geom_line() +
#   geom_point() +
#   xlab("Year") +
#   ylab("Parous (%)") +
#   ggtitle("Parity in England and Wales") +
#   scale_x_continuous(breaks = pretty_breaks())

# Plotting parity for birth cohorts

plot_data <- read_excel("Parity_Data.xlsx", sheet = "England_Wales_Parity") |>
  mutate(age = case_when(
    is.na(`Age exact years`) ~ 46,
    TRUE ~ `Age exact years`),
    `% Parous` = 100 - `0 (No children)`) |>
  group_by(`Year of birth of woman`) |>
  arrange(desc(age)) |>
  filter(row_number()==1) |>
  ungroup()

ggplot(plot_data, aes(x = `Year of birth of woman`, y = `% Parous`)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = max(plot_data$`Year of birth of woman`[plot_data$age==46])+0.5, linetype = "dashed", colour = "red") +
  xlab("Year of Birth") +
  labs(caption = "Dashed red line represents last full cohort (achieved age 46)")




# Plotting total parity in England/Wales over time.

pop_data <- read_excel("Population_Data.xlsx") |>
  filter(Sex == "Females", Age != "All Ages") |>
  select(-Sex) 

pop_data <- as.data.frame(t(as.matrix(pop_data))) 

colnames(pop_data) <- pop_data[1,]
pop_data$year <- rownames(pop_data)
pop_data <- pop_data[-1,] |>
  mutate(year = gsub(year, pattern = "Mid-", replacement = "")) |>
  sapply(as.numeric)

pop_data <- as.data.frame(pop_data) |>
  filter(year>1970)

for(i in 1:nrow(pop_data)){
  pop_data[i,which(colnames(pop_data)!="year")] <- pop_data[i,which(colnames(pop_data)!="year")]/sum(pop_data[i,which(colnames(pop_data)!="year")])
}

pop_data <- pop_data |>
  pivot_longer(cols = colnames(pop_data)[colnames(pop_data)!="year"])

parity_data <- read_excel("Parity_Data.xlsx", sheet = "England_Wales_Parity") |>
  mutate(flag =  case_when(
    is.na(`Age exact years`) ~ "Final",
    TRUE ~ NA),
    parous = 100 - `0 (No children)`,
    YOB = as.numeric(`Year of birth of woman`),
    age = as.numeric(`Age exact years`),
    year = YOB + age)

parity <- data.frame(year = integer(0), parity = integer(0))
  
for(y in 2010:2022){
  tempdata <- as.data.frame(parity_data)
  tempdata$finalage <- ifelse(parity_data$flag=="Final", y - tempdata$YOB, NA)
  tempdata <- filter(tempdata, year == y|finalage>45)
  tempdata$age <- ifelse(is.na(tempdata$age), tempdata$finalage, tempdata$age)
  over90 <- mean(tempdata$parous[tempdata$age>89])
  tempdata <- tempdata[tempdata$age<90,]
  tempdata <- select(tempdata, age, parous)
  tempdata[nrow(tempdata)+1,] <- c("90+", over90)
  tempdata$parous <- as.numeric(tempdata$parous)
  temppop <- pop_data[pop_data$year==y,]
  df <- merge(tempdata, temppop, by.x = "age", by.y = "name")
  df$relpar <- df$parous*df$value
  parity[nrow(parity)+1,] <- c(y, sum(df$relpar))
}

ggplot(parity, aes(x = year, y = parity)) +
  geom_point() +
  geom_line() +
  ylab("Parity (%)") +
  ggtitle("Parity in England and Wales") +
  scale_x_continuous(breaks = pretty_breaks())

```

## Fertility Rates

```{r}

fertility_data <- read_excel("Parity_Data.xlsx", sheet = "England_Wales_Fertility") |>
  filter(Year>1989)

ggplot(fertility_data, aes(x = Year, y = `Live births per 1,000 women in age group\r\nAll ages`)) +
  geom_point() +
  geom_line() +
  ylab("Fertility Rates per 1,000") +
  ggtitle("Fertility in England and Wales") +
  scale_x_continuous(breaks = pretty_breaks())

# heatmap_fertility <- fertility_data[,!grepl(colnames(fertility_data), pattern = "Age of mother")]
# 
# heatmap_fertility <- heatmap_fertility |>
#   pivot_longer(cols = colnames(heatmap_fertility)[-1]) |>
#   mutate(age_group = gsub(name, pattern = "Live births per 1,000 women in age group\r\n", replacement = "")) |>
#   filter(age_group != "All ages") |>
#   mutate(age_group = case_when(
#     grepl(age_group, pattern = "and over") ~ gsub(age_group, pattern = " and over", replacement = "+"),
#     grepl(age_group, pattern = "Under") ~ gsub(age_group, pattern = "Under ", replacement = "0 to "),
#     TRUE ~ age_group
#   ))
#   # group_by(age_group) |>
#   # mutate(value = (value - mean(value, na.rm = T))/sqrt(var(value, na.rm = T))) |>
#   # ungroup()
# 
# ggplot(heatmap_fertility, aes(x = age_group, y = Year, fill = value)) +
#   geom_tile() +
#   paletteer::scale_fill_paletteer_c("viridis::plasma") +
#   labs(fill = "Fertility Rate") +
#   xlab("Age Group") +
#   theme(legend.position = "bottom")


```

## HRT

```{r}

read_excel("HRT_Data.xlsx") |>
  mutate(Year = as.numeric(map_chr(strsplit(`Financial Year`, "/"), 1)), 
         `% HRT` = as.numeric(`Patients per 1,000 Population`)) |>
  filter(!is.na(`% HRT`)) |>
  ggplot(aes(x = Year, y = `% HRT`)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("HRT (%)") +
  ggtitle("HRT Usage in England") +
  scale_x_continuous(breaks = pretty_breaks()) 


```

## Cancer

```{r}

cancer_sheets <- excel_sheets("Cancer_Data.xlsx")

cancer_data <- integer(0)

for(sheet in cancer_sheets){
  cancer_data <- rbind(cancer_data, read_excel("Cancer_Data.xlsx", sheet = sheet))
}

cancer_data <- cancer_data |>
  mutate(Sex = ifelse(Sex==1, "Male", "Female"),
         IncMort = ifelse(Type == 0, "Incidence", "Mortality"))

# cancer_data |>
#   filter(Type == 0) |>
#   ggplot(aes(x = Year, y = `ASR (World)`, colour = `Country label`, linetype = Sex)) +
#   geom_line() + 
#   geom_point() +
#   ggtitle("Incidence Rates for Top 6 Cancers") +
#   facet_wrap(~ `Cancer label`, nrow = 3, scales = "free")
#   
# cancer_data |>
#   filter(Type == 1) |>
#   ggplot(aes(x = Year, y = `ASR (World)`, colour = `Country label`, linetype = Sex)) +
#   geom_line() + 
#   geom_point() +
#   ggtitle("Mortality Rates for Top 6 Cancers") +
#   facet_wrap(~ `Cancer label`, nrow = 3, scales = "free")

# for(site in unique(cancer_data$`Cancer label`)){
#   tempdata <- cancer_data[cancer_data$`Cancer label`==site,]
#   minyear <- min(tempdata$Year)
#   maxyear <- max(tempdata$Year)
#   print(ggplot(data = tempdata, aes(x = Year, y = `ASR (World)`, colour = `Country label`, linetype = Sex)) +
#   geom_line() +
#   geom_point() +
#   xlim(minyear, maxyear) +
#   ggtitle(paste0("Cancer Rates for ", site)) +
#   labs(colour = "Country", linetype = "Sex") +
#   theme(legend.position="bottom",legend.box = "vertical") +
#   facet_wrap(~ IncMort, scales = "free_y", dir = "v"))
# 
# }

for(site in unique(cancer_data$`Cancer label`)){
  tempdata <- cancer_data[cancer_data$`Cancer label`==site,]
  # minyear <- min(tempdata$Year)
  minyear <- 1989
  maxyear <- max(tempdata$Year)
  print(ggplot(data = tempdata, aes(x = Year, y = `ASR (World)`, colour = `Country label`)) +
  geom_line() +
  geom_point() +
  xlim(minyear, maxyear) +
  ggtitle(paste0("Cancer Rates for ", site)) +
  labs(colour = "Country") +
  theme(legend.position="bottom",legend.box = "vertical") +
  facet_grid(IncMort~ Sex, scales = "free_y"))

}

```

## Diabetes

```{r}

diabetes_data <- integer(0)

sheets <- excel_sheets("Diabetes_Data.xlsx")

england_sheets <- sheets[grepl(sheets, pattern = "England")]

for(sheet in england_sheets){

  nocol <- ncol(read_excel("Diabetes_Data.xlsx", sheet = sheet))
  tempdata <- read_excel("Diabetes_Data.xlsx", sheet = sheet, col_types = c("text", rep("numeric", nocol-1)))
  group <- gsub(x = sheet, pattern = "England_", replacement = "")
  tempdata <- pull_HSE_diabetes_data(data = tempdata, group = group)
  tempdata$section <- "Diabetes"
  diabetes_data <- rbind(diabetes_data, tempdata)
        
}

scotland_sheets <- sheets[grepl(sheets, pattern = "Scotland")]


for(sheet in scotland_sheets){
  tempdata <- read_excel("Diabetes_Data.xlsx", sheet = sheet)
  group <- gsub(x = sheet, pattern = "Scotland_", replacement = "")
  tempdata <- pull_SHS_data(data = tempdata, group = group)
  tempdata$section <- "Diabetes"
  tempdata$variable = ifelse(tempdata$variable == "Yes", "% Diabetes", "% No Diabetes")
  diabetes_data <- rbind(diabetes_data, tempdata)
}

tempdata <- read_excel("Diabetes_Data.xlsx", sheet = "NIreland") 

tempdata <- tempdata |>
  pivot_longer(cols = colnames(tempdata)[2:ncol(tempdata)]) |>
  rename(year = name, section = `...1`) |>
  mutate(value = value/10) |>
  mutate(age_group = "17+", variable = "% Diabetes", group = "All", country = "N. Ireland")

diabetes_data <- rbind(diabetes_data, tempdata)

diabetes_data |>
  mutate(year = case_when(
    grepl(year, pattern = "\\/") ~ map_chr(strsplit(year, "/"), 1),
    TRUE ~ year),
    year = as.numeric(year),
  age_group = case_when(
    age_group %in% c("All men", "All women", "All", "All adults", "Total") ~ "16+",
    grepl(age_group, pattern = "Age ") ~ gsub(age_group, pattern = "Age ", replacement = ""),
    grepl(age_group, pattern = " years") ~ gsub(age_group, pattern = " years", replacement = ""),
    grepl(age_group, pattern = "All adults ") ~ gsub(age_group, pattern = "All adults ", replacement = ""),
    grepl(age_group, pattern = "All aged ") ~ gsub(age_group, pattern = "All aged ", replacement = ""),
    TRUE ~ age_group
  )) |>
  filter(variable == "% Diabetes" & age_group %in% c("16+", "17+") & grepl(group, pattern = "All")) |>
  ggplot(aes(x = year, y = value, colour = country)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Diabetes") +
  ggtitle("% Diabetes in Adults") +
  labs(colour =  "Country", caption = "*Adult for England and Scotland defined as aged >15, for N. Ireland defined as aged >16.") +
  scale_x_continuous(breaks = pretty_breaks())


```

## Screening

```{r}

screening_data_breast <- read_excel("Screening_Data.xlsx", sheet = "England_Breast")

screening_data_breast |>
  mutate(Year = as.numeric(substr(`Time period`, start = 1, stop = 4))) |>
  ggplot(aes(x = Year, y = Value)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Uptake") +
  ggtitle("% Breast Screening Uptake in England") +
  scale_x_continuous(breaks = pretty_breaks())

screening_data_bowel <- read_excel("Screening_Data.xlsx", sheet = "England_Bowel")

screening_data_bowel |>
  mutate(Year = as.numeric(substr(`Time period`, start = 1, stop = 4))) |>
  ggplot(aes(x = Year, y = Value)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Coverage") +
  ggtitle("% Bowel Screening Coverage in England") +
  scale_x_continuous(breaks = pretty_breaks())

screening_data_cervix <- read_excel("Screening_Data.xlsx", sheet = "England_Cervix")

screening_data_cervix |>
  filter(...1 == "Coverage (25-64)") |>
  select(-units) |>
  pivot_longer(cols = colnames(screening_data_cervix)[substr(colnames(screening_data_cervix), start = 1, stop = 2) == "20"]) |>
  mutate(Year = as.numeric(name)) |>
  ggplot(aes(x = Year, y = value)) +
  geom_line() +
  geom_point() +
  xlab("Year") +
  ylab("% Coverage") +
  ggtitle("% Cervix Screening Coverage in England") +
  scale_x_continuous(breaks = pretty_breaks())
```
